The NIME of the Title Is Hello

### 2025-01-15 src/article.tex
Updates from Overleaf

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\FIGURE, but there was still a limit given the structural and aesthatic
requirements.

\subsubsection{Sensor Criteria}\label{sensor-criteria}

A set of criteria was drawn to which the final sensor system would have
to satisfy. This criteria was used as a means of evaluating the
feasability throughout the prototyping phase.

\begin{itemize}
\tightlist
\item
  Non-invasive: The sensor system should not require major modifications
  to the interface on which it is installed.
\item
  Low Latency: Reading sensors and then triggering an output should have
  a low latency. Using {[}paper on latency{]} a target was set as X ms
  with an upper limit set to 25ms beyond which the sensors would be
  deemed unusable
\item
  Reliable: Data obtained should be reliable and accurate enough not to
  result in false positives or false negatives without extensive
  filtering
\item
  Scalable: The system should be scalable both economically and
  temporally. Given the open soucre nature of the porject, it could not
  be prohibitvely expensive to carry out. Also any system that would
  work on a single key would have to scale financially to potentially 50
  or more. The time required to install and calibrate the sensors would
  also have to scale.
\item
  Expandable: Any sensor system should be functionally expandable to
  allow for the usage of the data for control over other parameters.
\end{itemize}

\subsubsection{Triggering a note}\label{triggering-a-note}

For the first iteration the data from the sensor would simply be used to
test when a threshold had been passed. On passing the threshold a
message would be sent to trigger the playback of a corresponding note.

The message format covered in \textbackslash section

\section{Related Work and Motivation}\label{related-work-and-motivation}

\begin{itemize}
\tightlist
\item
  \cite{HapticKey}
\item
  \cite{mcPherson1} \cite{mcPherson2}
\item
  \cite{Mudd2013}
\item
  \cite{FritzClaudia2017Leon}
\end{itemize}

\section{Hardware Design}\label{hardware-design}

The hardware was designed with the xx in mind.

Using a similalr apporach to @HapticKey the first step was to fabricate
or procure a model of a single working key. San Colombano provided a
model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
3-key model was used a test-bed for potential sensors while designs were
drawn for a full-scale model. The cross-sectional nature of the model
meant it was easy to fit and remove sensors. Small size meant that it
would also be easy to transfer between workshops.

3-key model also had a sliding jack rail which allowed for easy
disengagement of the physical mechanical response, which allowed for
some preliminary tests on the presence and absence of mechanism.

\subsection{Failed sensors}\label{failed-sensors}

A number of sensors were tried before the final system was decided upon.
An inertial measurment unit (IMU) was mounted to a single key which
contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
a consistent set of conditions to trigger sound, but this was not easily
differntiable when other keys were played. The fitting time and cost of
a single IMU could b reduced when purchased in bulk but was far greater
than other systems considered.

A reed sensor and a hall sensor were mounted and magnets embedded within
the jacks. Adjusting the hieight of the sensor provided a reliable means
of detecting a threshold, inparticular the hall sensors which could be
partially calibrarted physically and also in softwrae. The time taken to
calibarte a single key was time consuming. Ideally each sensor would be
placed on a thread for adjustment, but given the constraints on the
internal space it was decided any adjustment of heights would likely
mean an unnacceptable length of time to install and calibrate. Embedding
magnets into the jacks was also did not satisfy the criteria of being
scalable and non-invasive.

Light dependant resistor mounted to the jack rail. as jacks were pressed
less light would be let through. Usable data but very dependant on
lighting conditions. Modifying the jack also did not satisfy the
non-intrusive criteria and the placement outside the internal chamber of
the model meant hiding electronics would be more difficult.

Finally, force sensitive resistors placed under the jack of each key was
considered but was never implemented. The sensors were to be placed
under each jack either above or below the padding on which the bottom of
the jack rests. Further discussion of this apporach is in
\textbackslash section.

\section{Electronics}\label{electronics}

Design of electronics carried out in tandem with construction of full
model. Since the Electronics were designed at a distanced to the final
full model some flexbility was required in the design. Design decisions
that were not as optimal but would allow for change in wiring and
fitting should there be any issues during installation.

Electronics designed so that they would not require any equipment
outside what can commonly be found in a maker space at a univeristy.

The equipment required including soldering suitable for throughole
technology (THT) components and 3D printing.

Laser cutting and CNC routing fascilities were used during the research
for the project but not required for the final implementation.

The idea behind this was to allow for the project to be reasonably
recreated without procurement of any additional specialist equipment.

NEMUS lab under construction during the first section of the project.
Facilities available at the maker spaces of Bologna (Alma Labor) and
Edinburgh (uCreate) used for what was deemed reasonable assumption in
terms of available equipment.

\begin{itemize}
\tightlist
\item
  Over arching themes

  \begin{itemize}
  \tightlist
  \item
    simplicity of design

    \begin{itemize}
    \tightlist
    \item
      Limited resources
    \item
      easily reproducible
    \end{itemize}
  \item
    availability and lead times
  \item
    iteration and design from distance
  \end{itemize}
\end{itemize}

\subsubsection{QRE1113}\label{qre1113}

Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
combination infrared LED and phototransistor sensitive to IR light. The
phototransistor of QRE1113 provides data on how much light is present
and in particular how much light is being relfected from a surface.
Since refelected light will be proportional to the distance of a surfcae
it is a good, close proximity, distance sensor. Distance is the way in
which the instruments in \citeMcpherson1\&2 function as well the Moog
Piano Bar from which it too inspiration.

The limitation for taking a distance approach is that there is not one
but 2 jacks per key. For each jack to be measured indepeedntly the data
would need tp com fropm the jacks themselves.

They \emph{could} be used and individual trigger points set in
calibration, but this creates new problems.

\begin{itemize}
\tightlist
\item
  The triggering would need to be bespoke for every key as fluxuations
  in the dimensions of every piece mean that the threshold would be
  different.

  \begin{itemize}
  \tightlist
  \item
    Moving a problem into additional time in the calibtraryoin setup
    phase.
  \end{itemize}
\item
  Remebering the constraint that system isshould noyt fix problem, there
  would be no easy way to tell if a jack had nbot reseated.

  \begin{itemize}
  \tightlist
  \item
    In playing a harpihchord when the jack does not go to its rest
    position, not note would play. This fault should be carried over
    intp the final instrument.
  \end{itemize}
\end{itemize}

Another common application of the QRE1113 is within line following
robots, where an array of QRE1113s ar used in conjunction with black
tape along which some motorised system can following

Applied to the jacks was a greyscale gradient printed the length full
travel range of the jack.

For the first iteration this was performed on an inkjet printer and
double-sided tape. For the final iteration the stickers were printed on
(Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
reduced the manual labour for cutting stickers meaning the process would
scale from the 3-key to the 49-key model.

An initial sticker design was used that contained only the gradient. It
was found that the jacks wer sticking in their slots andthe edges on the
stickers easily peeled around the printing The sticker was redesigned to
be the full length of the jack which avoided the problem of peeling. The
larger are also mean the stickers adhered betterto the jack Some
trimming was still required of each sticker to avoid them adding extraa
resistence to the jack's movement.

During testing it was found that their was a considerable amount
cross-talk bwteen adjacent sensors. This meant the reading from key
would be affected by the movement of the key next to it. The cross-talk
was assumed to be as a result of light reflecting from other jacks.
Given that IR is used this was difficult to confirm by sight. A set of
baffles was designed for the pcb that would wrap round the jack and stop
light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
printers using PLA plastic. A dark pigmented PLA was used to avoid IR
light simply being diffused. After the baffles were fitted the
cross-talk was entirely eliminated.

\begin{itemize}
\tightlist
\item
  previous use
\item
  problem of what to sense
\item
  satisfied both criteria and design restrictions
\item
  usage

  \begin{itemize}
  \tightlist
  \item
    Sweet spot is around 6mm distance
  \end{itemize}
\end{itemize}

\subsubsection{sensor surface}\label{sensor-surface}

\begin{itemize}
\tightlist
\item
  linear gradient
\item
  on Jack
\item
  printed and cut on vinyl {[}find name{]}
\item
  cut to travel size
\item
  then to jack size
\item
  jacks varnished on side
\item
  alternatives

  \begin{itemize}
  \tightlist
  \item
    CNC plywood and mdf

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \item
    laser cut plywood and mdf

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \item
    3D printed jack body

    \begin{itemize}
    \tightlist
    \item
      problems
    \end{itemize}
  \end{itemize}
\end{itemize}

\subsection{Arduino}\label{arduino}

\subsubsection{Nano BLE}\label{nano-ble}

Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
appealing as it would allow for easy changes between chipsets without
requiring any rewiring and thus no new PCBs would need to be created if
a better alternative was found later in the project. Another benefit to
the Arduino Nano boards was that typically had native-USB meaning the
could be programmed as a hardware USB MIDI device directly.

The Nano BLE specifically was used in the initial stages for the
integrated IMU.

When it was finalised that the prokect would proceed using the QRE1113 a
number of microcontroller units (MCU) were condsidered.

The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
Arduino Nano ESP32 and STM32 NUCLEO-L031K6.

A rough performance test was later carried out using an STM32
NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
more performative. To execute a single sensor read cycle of the firmware
the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
on average. The Nano ESP32 had high variability so it was decided to
continue with the Nano BLE. Further discussion of using ESP32 continues
in

\section{Going forward}

BLE functionality also meant if the cable connection between the nano
and computer running audio synthesis was impractical, BLE MIDI was
available as a fallback. The Nano BLE also has a 12-bit DAC, which
provided a contingency should the default 10-bit not be sensitive enough
for the signal from the sensors.

Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
functionality, but it was found that 2 ADC channels were in fact not
useable and would have required additional multiplexers.

\begin{itemize}
\tightlist
\item
  Nano form factor small
\item
  allows for easy changes in chipset
\item
  ``Native USB''
  https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
\item
  BLE Model

  \begin{itemize}
  \tightlist
  \item
    Performative
  \item
    BLE capabilities for MIDI BLE transmission
  \item
    All analog pins available
  \item
    available at the time
  \item
    12-bit DAC
  \end{itemize}
\end{itemize}

\subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}

There would need to be the ability to store values on non-volatile
memory so that the electroincs could be powered-down without losing
data.

Adjusting the values through firmware would not have been flexible or
practical. Instead an Ferroelectric RAM unit was chosen and connected
using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
was primarily chosen for it's ready availability.

Most microcontrollers have some variety of non-volatile program memory
that can be read and writ accessed but in most cases this memory is
cleared during compilation and re-flashing. The small cost was far
outweighed by the potential for data loss if using onboard RAM.

\emph{It is worth noting that the FRAM model chosen stopped being
available during the project, meaning any future iteration would need to
adjust for another board in firmware}

The firmware was structure in a way that the interface could be changed
so long as the code required for any other board would comply with
function arguments and outputs. Worth noting the fragility and
volatility when designing this kind of project and how easily it is for
the project to become broken.

SD card would provide an easier way to interact with calibration data.
The posiiton of the controller board did not permit easy access.
reliability and speed was preffered over storage capacity.

\begin{itemize}
\tightlist
\item
  On board available, but volatile
\item
  retained across firmware changes
\item
  small,
\item
  could have been sd card or anything else
\item
  available
\item
  interfaced via rotary encoder
\end{itemize}

\subsection{RGB LEDs}\label{rgb-leds}

for problems of scale:

Addressing the problem of scale of identifying which key's threshold is
currently being edited.

The LEDs also allowed for easy identification for keys that were outside
of expected values and when a key was triggered. This allowed for visual
identification of malfunctioning or uncalibrated keys.

The LEDs are on the same PCB as the sensors and are hidden when the
nameboard is in place.

Would sugggest that this is a necessity of this kind to provide a means
to identify and diagnose LEDs with state.

TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
to assembling, but were avoid in favour of Addressable LEDs with
intergrated driver. The intergrated LEDs share the same data line which
has the limitation where a fualty LED will stop all following LEDs from
working. An individual PWM driver per PCB would allow avoid situations
where failutre cascades, but would add complexity in assembly as the
parts are typically SMD. Also, given the LEDs are fixed to the board and
could not be repaired in situ, the solution would be to replace the PCB
regardless of which approach was taken. Given the LEDs were only used in
debugging and calibration and otehrwise would not be visible, a simpler
design approach was taken.

\begin{itemize}
\tightlist
\item
  locating keys
\item
  displaying key state
\end{itemize}

\subsection{Multiplexing}\label{multiplexing}

CD4051BE Texas Instruments

problem of scale

For the 3-key model, individual jacks could be connected to one of the 8
available ADC pins.

The BLE Nano has 8 ADC pins connected to one ADC channel via a
multiplexer

Multiplexer added to expand to the

Transistor / Op-amp circuit could have been used similar to the one in
\cite{mcpherson}.

Benefits were keeping with THT based designed for easy assembly.

An 8-channel multiplexer was chosen as the model has 49-keys and a
division of 7 PCBs each containing 7 sensors

The multiplexer reduced 7 signals to a single pin. Three pins are
required pins for addressing the muliplexewrs and all PCBs are addressed
simulatenously.

\begin{itemize}
\tightlist
\item
  limited ADC channels
\item
  ADC channels already multiplexed
\item
  simple
\item
  8 channel
\item
  THT
\item
  7 into 49

  \begin{itemize}
  \tightlist
  \item
    7 PCBS
  \item
    each with 7 sensors
  \end{itemize}
\end{itemize}

\subsection{PCB design}\label{pcb-design}

\begin{itemize}
\tightlist
\item
  EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
  deprecated EAGLE and the programme will no longer be supported from
  2026 Project files will need to be transitioned to another standard,
  likely the KiCAD programme.
\end{itemize}

trade off between limiting number of boards required and allowing
compensation for variation in pitch.

Smaller PCBs means that spacing could be adjusted to accomodate the
changes in pitch between jacks. It also allowed for a modular approach
where any PCB could swapped for minimal cost.

\subsubsection{Sensor board}\label{sensor-board}

The 49 QRE1113s are divided across 7 PCBs each containing

\begin{itemize}
\item
  x7 QRE1113
\item
  x7 100R
\item
  x7 10kR
\item
  x1 CD4051BE Multipexer
\item
  jack / sensor pitch by board
\item
  7 boards adjust for pitch/spacing error
\item
  modular, replaceable
\item
  orientation problems

  \begin{itemize}
  \tightlist
  \item
    legs break easily
  \item
    end mount like A.P.Mc. time consuming
  \item
    problem

    \begin{itemize}
    \tightlist
    \item
      milling out keys
    \end{itemize}
  \end{itemize}
\item
  designed from distance
\item
  striped lines across pcbs
\item
  problems

  \begin{itemize}
  \tightlist
  \item
    additional terminals uneeded resulted in lots of additional work
  \item
    long cabling increase risk of breaking
  \end{itemize}
\item
  solutions

  \begin{itemize}
  \tightlist
  \item
    striping all adc lines with FFC
  \item
    solder pads to select channel
  \end{itemize}
\end{itemize}

\subsubsection{controller board}\label{controller-board}

Controller board designed around Arduino nano form factor. The Nano is
connected through headers to allow easy switching between chips. Should
another chip in a Nano for factor be found it could supplant the current
Nano without any need for a redesign.

The first version of the board esclusively used 2.54mm pitch headers.
The reasons for this were flexibility when creating cable looms and to
leave open the possiility of rewuring the PCB.

This was the case for the LEDs as they initially wer to be powered on
5V. It was discovered that the 3v regulator of the nano was sufficient
to power all LEDs at a lower brightness while reducing current draw.
Since brightness was not importnat.

The 2.54mm pitch headers had two major drawbacks. First, the orientation
of the cables could be reversed, which in the best case scenario would
simply mean the system did not work correctly and in the worst case it
could potentially damaged some components.

Secondly, the headers do not connect securely and wait of the cabling
was enough for components to become disconnected as the harpsichord was
handled.

To address this, once the wiring for the controller board was finalised
a new design was drafted this time using JST-PH connectors. No
commercial cables were immediately available in the configuration and
length required, as such a cable loom needed to be made which meant
crimping cables by hand. A crimping tool from JST is prohibitively
expensive compared to the cost of other components in the project*. The
crimping tool also appeared to be uncommon and was not readily available
in the workshops at the projects disposal. Eventually a tool was found,
but an alternative approach was also researched in order to maintain the
criteria of accessibility.

A locator was adapted from a creative commons stl model
https://www.thingiverse.com/thing:1646016 that allowed a generic
crimping tool to be adapted for use with JST-PH crimps† \figure{}. The
locator was printed using a Bambu X1 3D printer.

>>>>

Accomodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the eletronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\\FIGURE, but there was still a limit given the structural and aesthatic
requirements.

% \subsubsection{Sensor Criteria}\label{sensor-criteria}

% A set of criteria was drawn to which the final sensor system would have
% to satisfy. This criteria was used as a means of evaluating the
% feasability throughout the prototyping phase.

% \begin{itemize}
% \tightlist
% \item
%   Non-invasive: The sensor system should not require major modifications
%   to the interface on which it is installed.
% \item
%   Low Latency: Reading sensors and then triggering an output should have
%   a low latency. Using {[}paper on latency{]} a target was set as X ms
%   with an upper limit set to 25ms beyond which the sensors would be
%   deemed unusable
% \item
%   Reliable: Data obtained should be reliable and accurate enough not to
%   result in false positives or false negatives without extensive
%   filtering
% \item
%   Scalable: The system should be scalable both economically and
%   temporally. Given the open soucre nature of the porject, it could not
%   be prohibitvely expensive to carry out. Also any system that would
%   work on a single key would have to scale financially to potentially 50
%   or more. The time required to install and calibrate the sensors would
%   also have to scale.
% \item
%   Expandable: Any sensor system should be functionally expandable to
%   allow for the usage of the data for control over other parameters.
% \end{itemize}

% \subsubsection{Triggering a note}\label{triggering-a-note}

% For the first iteration the data from the sensor would simply be used to
% test when a threshold had been passed. On passing the threshold a
% message would be sent to trigger the playback of a corresponding note.

% The message format covered in \textbackslash section

% \section{Related Work and Motivation}\label{related-work-and-motivation}

% \begin{itemize}
% \tightlist
% \item
%   \cite{HapticKey}
% \item
%   \cite{mcPherson1} \cite{mcPherson2}
% \item
%   \cite{Mudd2013}
% \item
%   \cite{FritzClaudia2017Leon}
% \end{itemize}

% \section{Hardware Design}\label{hardware-design}

% The hardware was designed with the xx in mind.

% Using a similalr apporach to @HapticKey the first step was to fabricate
% or procure a model of a single working key. San Colombano provided a
% model 3-key of a harposchord mechanism by Graziano Bandini \FIGURE. The
% 3-key model was used a test-bed for potential sensors while designs were
% drawn for a full-scale model. The cross-sectional nature of the model
% meant it was easy to fit and remove sensors. Small size meant that it
% would also be easy to transfer between workshops.

% 3-key model also had a sliding jack rail which allowed for easy
% disengagement of the physical mechanical response, which allowed for
% some preliminary tests on the presence and absence of mechanism.

% \subsection{Failed sensors}\label{failed-sensors}

% A number of sensors were tried before the final system was decided upon.
% An inertial measurment unit (IMU) was mounted to a single key which
% contained a 6 degrees of freedome acceleromoetr and gyroscope. There was
% a consistent set of conditions to trigger sound, but this was not easily
% differntiable when other keys were played. The fitting time and cost of
% a single IMU could b reduced when purchased in bulk but was far greater
% than other systems considered.

% A reed sensor and a hall sensor were mounted and magnets embedded within
% the jacks. Adjusting the hieight of the sensor provided a reliable means
% of detecting a threshold, inparticular the hall sensors which could be
% partially calibrarted physically and also in softwrae. The time taken to
% calibarte a single key was time consuming. Ideally each sensor would be
% placed on a thread for adjustment, but given the constraints on the
% internal space it was decided any adjustment of heights would likely
% mean an unnacceptable length of time to install and calibrate. Embedding
% magnets into the jacks was also did not satisfy the criteria of being
% scalable and non-invasive.

% Light dependant resistor mounted to the jack rail. as jacks were pressed
% less light would be let through. Usable data but very dependant on
% lighting conditions. Modifying the jack also did not satisfy the
% non-intrusive criteria and the placement outside the internal chamber of
% the model meant hiding electronics would be more difficult.

% Finally, force sensitive resistors placed under the jack of each key was
% considered but was never implemented. The sensors were to be placed
% under each jack either above or below the padding on which the bottom of
% the jack rests. Further discussion of this apporach is in
% \textbackslash section.

% \section{Electronics}\label{electronics}

% Design of electronics carried out in tandem with construction of full
% model. Since the Electronics were designed at a distanced to the final
% full model some flexbility was required in the design. Design decisions
% that were not as optimal but would allow for change in wiring and
% fitting should there be any issues during installation.

% Electronics designed so that they would not require any equipment
% outside what can commonly be found in a maker space at a univeristy.

% The equipment required including soldering suitable for throughole
% technology (THT) components and 3D printing.

% Laser cutting and CNC routing fascilities were used during the research
% for the project but not required for the final implementation.

% The idea behind this was to allow for the project to be reasonably
% recreated without procurement of any additional specialist equipment.

% NEMUS lab under construction during the first section of the project.
% Facilities available at the maker spaces of Bologna (Alma Labor) and
% Edinburgh (uCreate) used for what was deemed reasonable assumption in
% terms of available equipment.

% \begin{itemize}
% \tightlist
% \item
%   Over arching themes

%   \begin{itemize}
%   \tightlist
%   \item
%     simplicity of design

%     \begin{itemize}
%     \tightlist
%     \item
%       Limited resources
%     \item
%       easily reproducible
%     \end{itemize}
%   \item
%     availability and lead times
%   \item
%     iteration and design from distance
%   \end{itemize}
% \end{itemize}

% \subsubsection{QRE1113}\label{qre1113}

% Following from the work in \cite{McPherson1} \cite{McPherson2} we tested
% a system using the Fairchild QRE1113 \datasheetfigue. The QRE1113 is a
% combination infrared LED and phototransistor sensitive to IR light. The
% phototransistor of QRE1113 provides data on how much light is present
% and in particular how much light is being relfected from a surface.
% Since refelected light will be proportional to the distance of a surfcae
% it is a good, close proximity, distance sensor. Distance is the way in
% which the instruments in \citeMcpherson1\&2 function as well the Moog
% Piano Bar from which it too inspiration.

% The limitation for taking a distance approach is that there is not one
% but 2 jacks per key. For each jack to be measured indepeedntly the data
% would need tp com fropm the jacks themselves.

% They \emph{could} be used and individual trigger points set in
% calibration, but this creates new problems.

% \begin{itemize}
% \tightlist
% \item
%   The triggering would need to be bespoke for every key as fluxuations
%   in the dimensions of every piece mean that the threshold would be
%   different.

%   \begin{itemize}
%   \tightlist
%   \item
%     Moving a problem into additional time in the calibtraryoin setup
%     phase.
%   \end{itemize}
% \item
%   Remebering the constraint that system isshould noyt fix problem, there
%   would be no easy way to tell if a jack had nbot reseated.

%   \begin{itemize}
%   \tightlist
%   \item
%     In playing a harpihchord when the jack does not go to its rest
%     position, not note would play. This fault should be carried over
%     intp the final instrument.
%   \end{itemize}
% \end{itemize}

% Another common application of the QRE1113 is within line following
% robots, where an array of QRE1113s ar used in conjunction with black
% tape along which some motorised system can following

% Applied to the jacks was a greyscale gradient printed the length full
% travel range of the jack.

% For the first iteration this was performed on an inkjet printer and
% double-sided tape. For the final iteration the stickers were printed on
% (Coala, 1D 100 Gloss P, gloss, white, permanent adhesive, 300g/m2, 100
% µm, 1370mm x 50.00m, Solvent/Latex/UV) using a HP Latex 115 and then
% cutout with a Summa 150 roll cutter. The printer and roll cutter greatly
% reduced the manual labour for cutting stickers meaning the process would
% scale from the 3-key to the 49-key model.

% An initial sticker design was used that contained only the gradient. It
% was found that the jacks wer sticking in their slots andthe edges on the
% stickers easily peeled around the printing The sticker was redesigned to
% be the full length of the jack which avoided the problem of peeling. The
% larger are also mean the stickers adhered betterto the jack Some
% trimming was still required of each sticker to avoid them adding extraa
% resistence to the jack's movement.

% During testing it was found that their was a considerable amount
% cross-talk bwteen adjacent sensors. This meant the reading from key
% would be affected by the movement of the key next to it. The cross-talk
% was assumed to be as a result of light reflecting from other jacks.
% Given that IR is used this was difficult to confirm by sight. A set of
% baffles was designed for the pcb that would wrap round the jack and stop
% light from other jacks or sensors. Baffles were printed on BAMBU X1 3D
% printers using PLA plastic. A dark pigmented PLA was used to avoid IR
% light simply being diffused. After the baffles were fitted the
% cross-talk was entirely eliminated.

% \begin{itemize}
% \tightlist
% \item
%   previous use
% \item
%   problem of what to sense
% \item
%   satisfied both criteria and design restrictions
% \item
%   usage

%   \begin{itemize}
%   \tightlist
%   \item
%     Sweet spot is around 6mm distance
%   \end{itemize}
% \end{itemize}

% \subsubsection{sensor surface}\label{sensor-surface}

% \begin{itemize}
% \tightlist
% \item
%   linear gradient
% \item
%   on Jack
% \item
%   printed and cut on vinyl {[}find name{]}
% \item
%   cut to travel size
% \item
%   then to jack size
% \item
%   jacks varnished on side
% \item
%   alternatives

%   \begin{itemize}
%   \tightlist
%   \item
%     CNC plywood and mdf

%     \begin{itemize}
%     \tightlist
%     \item
%       problems
%     \end{itemize}
%   \item
%     laser cut plywood and mdf

%     \begin{itemize}
%     \tightlist
%     \item
%       problems
%     \end{itemize}
%   \item
%     3D printed jack body

%     \begin{itemize}
%     \tightlist
%     \item
%       problems
%     \end{itemize}
%   \end{itemize}
% \end{itemize}

% \subsection{Arduino}\label{arduino}

% \subsubsection{Nano BLE}\label{nano-ble}

% Arduino Nano BLE was chosen for a few reasons. The Nano form factor was
% appealing as it would allow for easy changes between chipsets without
% requiring any rewiring and thus no new PCBs would need to be created if
% a better alternative was found later in the project. Another benefit to
% the Arduino Nano boards was that typically had native-USB meaning the
% could be programmed as a hardware USB MIDI device directly.

% The Nano BLE specifically was used in the initial stages for the
% integrated IMU.

% When it was finalised that the prokect would proceed using the QRE1113 a
% number of microcontroller units (MCU) were condsidered.

% The MCUs considered were the Arduino Nano 33 IoT, Arduino Nano 33 BLE,
% Arduino Nano ESP32 and STM32 NUCLEO-L031K6.

% A rough performance test was later carried out using an STM32
% NUCLEO-L031K6, a Nano ESP32 and the BLE. The Nano BLE was found to be
% more performative. To execute a single sensor read cycle of the firmware
% the STM32 took 40ms, and both the Nano ESP32 and Nano BLE required 11ms
% on average. The Nano ESP32 had high variability so it was decided to
% continue with the Nano BLE. Further discussion of using ESP32 continues
% in

% \section{Going forward}

% BLE functionality also meant if the cable connection between the nano
% and computer running audio synthesis was impractical, BLE MIDI was
% available as a fallback. The Nano BLE also has a 12-bit DAC, which
% provided a contingency should the default 10-bit not be sensitive enough
% for the signal from the sensors.

% Initially a Nano IoT was used for both WiFi and Bluetooth Low Energy
% functionality, but it was found that 2 ADC channels were in fact not
% useable and would have required additional multiplexers.

% \begin{itemize}
% \tightlist
% \item
%   Nano form factor small
% \item
%   allows for easy changes in chipset
% \item
%   ``Native USB''
%   https://learn.adafruit.com/dude-where-s-my-com-port/native-usb-boards\#
% \item
%   BLE Model

%   \begin{itemize}
%   \tightlist
%   \item
%     Performative
%   \item
%     BLE capabilities for MIDI BLE transmission
%   \item
%     All analog pins available
%   \item
%     available at the time
%   \item
%     12-bit DAC
%   \end{itemize}
% \end{itemize}

% \subsection{EEPROM / NVRAM / FRAM}\label{eeprom-nvram-fram}

% There would need to be the ability to store values on non-volatile
% memory so that the electroincs could be powered-down without losing
% data.

% Adjusting the values through firmware would not have been flexible or
% practical. Instead an Ferroelectric RAM unit was chosen and connected
% using an SPI interface. NVRAM or EEPROM would have sufficed and the FRAM
% was primarily chosen for it's ready availability.

% Most microcontrollers have some variety of non-volatile program memory
% that can be read and writ accessed but in most cases this memory is
% cleared during compilation and re-flashing. The small cost was far
% outweighed by the potential for data loss if using onboard RAM.

% \emph{It is worth noting that the FRAM model chosen stopped being
% available during the project, meaning any future iteration would need to
% adjust for another board in firmware}

% The firmware was structure in a way that the interface could be changed
% so long as the code required for any other board would comply with
% function arguments and outputs. Worth noting the fragility and
% volatility when designing this kind of project and how easily it is for
% the project to become broken.

% SD card would provide an easier way to interact with calibration data.
% The posiiton of the controller board did not permit easy access.
% reliability and speed was preffered over storage capacity.

% \begin{itemize}
% \tightlist
% \item
%   On board available, but volatile
% \item
%   retained across firmware changes
% \item
%   small,
% \item
%   could have been sd card or anything else
% \item
%   available
% \item
%   interfaced via rotary encoder
% \end{itemize}

% \subsection{RGB LEDs}\label{rgb-leds}

% for problems of scale:

% Addressing the problem of scale of identifying which key's threshold is
% currently being edited.

% The LEDs also allowed for easy identification for keys that were outside
% of expected values and when a key was triggered. This allowed for visual
% identification of malfunctioning or uncalibrated keys.

% The LEDs are on the same PCB as the sensors and are hidden when the
% nameboard is in place.

% Would sugggest that this is a necessity of this kind to provide a means
% to identify and diagnose LEDs with state.

% TLC5940 or other PWM driver such as in \cite{mcpherson} added complexity
% to assembling, but were avoid in favour of Addressable LEDs with
% intergrated driver. The intergrated LEDs share the same data line which
% has the limitation where a fualty LED will stop all following LEDs from
% working. An individual PWM driver per PCB would allow avoid situations
% where failutre cascades, but would add complexity in assembly as the
% parts are typically SMD. Also, given the LEDs are fixed to the board and
% could not be repaired in situ, the solution would be to replace the PCB
% regardless of which approach was taken. Given the LEDs were only used in
% debugging and calibration and otehrwise would not be visible, a simpler
% design approach was taken.

% \begin{itemize}
% \tightlist
% \item
%   locating keys
% \item
%   displaying key state
% \end{itemize}

% \subsection{Multiplexing}\label{multiplexing}

% CD4051BE Texas Instruments

% problem of scale

% For the 3-key model, individual jacks could be connected to one of the 8
% available ADC pins.

% The BLE Nano has 8 ADC pins connected to one ADC channel via a
% multiplexer

% Multiplexer added to expand to the

% Transistor / Op-amp circuit could have been used similar to the one in
% \cite{mcpherson}.

% Benefits were keeping with THT based designed for easy assembly.

% An 8-channel multiplexer was chosen as the model has 49-keys and a
% division of 7 PCBs each containing 7 sensors

% The multiplexer reduced 7 signals to a single pin. Three pins are
% required pins for addressing the muliplexewrs and all PCBs are addressed
% simulatenously.

% \begin{itemize}
% \tightlist
% \item
%   limited ADC channels
% \item
%   ADC channels already multiplexed
% \item
%   simple
% \item
%   8 channel
% \item
%   THT
% \item
%   7 into 49

%   \begin{itemize}
%   \tightlist
%   \item
%     7 PCBS
%   \item
%     each with 7 sensors
%   \end{itemize}
% \end{itemize}

% \subsection{PCB design}\label{pcb-design}

% \begin{itemize}
% \tightlist
% \item
%   EAGLE PCBs designed using EAGLE. Worth noting that AutoCAD has
%   deprecated EAGLE and the programme will no longer be supported from
%   2026 Project files will need to be transitioned to another standard,
%   likely the KiCAD programme.
% \end{itemize}

% trade off between limiting number of boards required and allowing
% compensation for variation in pitch.

% Smaller PCBs means that spacing could be adjusted to accomodate the
% changes in pitch between jacks. It also allowed for a modular approach
% where any PCB could swapped for minimal cost.

% \subsubsection{Sensor board}\label{sensor-board}

% The 49 QRE1113s are divided across 7 PCBs each containing

% \begin{itemize}
% \item
%   x7 QRE1113
% \item
%   x7 100R
% \item
%   x7 10kR
% \item
%   x1 CD4051BE Multipexer
% \item
%   jack / sensor pitch by board
% \item
%   7 boards adjust for pitch/spacing error
% \item
%   modular, replaceable
% \item
%   orientation problems

%   \begin{itemize}
%   \tightlist
%   \item
%     legs break easily
%   \item
%     end mount like A.P.Mc. time consuming
%   \item
%     problem

%     \begin{itemize}
%     \tightlist
%     \item
%       milling out keys
%     \end{itemize}
%   \end{itemize}
% \item
%   designed from distance
% \item
%   striped lines across pcbs
% \item
%   problems

%   \begin{itemize}
%   \tightlist
%   \item
%     additional terminals uneeded resulted in lots of additional work
%   \item
%     long cabling increase risk of breaking
%   \end{itemize}
% \item
%   solutions

%   \begin{itemize}
%   \tightlist
%   \item
%     striping all adc lines with FFC
%   \item
%     solder pads to select channel
%   \end{itemize}
% \end{itemize}

% \subsubsection{controller board}\label{controller-board}

% Controller board designed around Arduino nano form factor. The Nano is
% connected through headers to allow easy switching between chips. Should
% another chip in a Nano for factor be found it could supplant the current
% Nano without any need for a redesign.

% The first version of the board esclusively used 2.54mm pitch headers.
% The reasons for this were flexibility when creating cable looms and to
% leave open the possiility of rewuring the PCB.

% This was the case for the LEDs as they initially wer to be powered on
% 5V. It was discovered that the 3v regulator of the nano was sufficient
% to power all LEDs at a lower brightness while reducing current draw.
% Since brightness was not importnat.

% The 2.54mm pitch headers had two major drawbacks. First, the orientation
% of the cables could be reversed, which in the best case scenario would
% simply mean the system did not work correctly and in the worst case it
% could potentially damaged some components.
